---
title: "DATA 605 - Homework 15"
author: "Manolis Manoli"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---
```{r loading-libraries, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(psych)
library(dplyr)
library(knitr)
library(tidyr)
```

\section{1. Regression Line}

Find the equation of the regression line for the given points.  Round any final values to the nearest hundredth, if necessary.( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

```{r }
regdata <- matrix(c( 6.3, 12.4 ,  7, 14.8 ,  7.7, 18.2 ,  8.4, 20.8 ),ncol=2,byrow=TRUE)
colnames(regdata) <- c("X","Y")
regdf = as.data.frame(regdata)
```


```{r }
regr = lm(regdf$Y ~ regdf$X)
summary(regr)
```

```{r }
ggplot(regdf, aes(x = X, y = Y)) +
geom_point() +
geom_smooth(method='lm', formula= y ~ x )
```

```{r }
inter = summary(regr)$coefficients[1,1]
slope = summary(regr)$coefficients[2,1]
```

The fitted line has the formula: y = `r round(inter,3)` + `r round(slope,3)` * x

For each function, only consider its valid ranges as indicated in the notes when you are computing the Taylor Series expansion. 

\section{2. Curve Discussion}

Find all local maxima, local minima, and saddle points for the function given below.  Write your answer(s) in the form (x, y, z).  Separate multiple points with a comma. $f(x, y) = 24x - 6xy^2 - 8y^3$

$f(x, y) = 24x - 6xy^2 - 8y^3$

$f_x = 24 - 6y^2$

$f_y = -12xy - 24y^2$

$f_{xx} = 0$

$f_{xy} = -12y$

$f_{yx} = -12y$

$f_{yy} = -12x - 48y$

find where the first derivatives are zero:

$f_x = 24 - 6y^2 = 0$ <=> $6y^2 = 24$  <=> $y^2 = 4$ <=> $y = \pm 2$

we can plug the two solutions into $f_y$.

$y=2$:

$f_y = -24x - 96 = 0$ <=> $24x = -96$  <=> $x = -4$

First point is (-4,2)

$f(-4,2)= -64$

 (-4,2,-64)

$y=-2$:

$f_y = 24x - 96 = 0$ <=> $24x = 96$  <=> $x = 4$

Second point is (4,-2)

$f(4,-2)= 64$

(4,-2,64)

we need to calculate $D=f_{xx}f_{yy}−f_{xy}^2$

Since $f_{xx} = 0$ $D=−f_{xy}^2 = -(-12y)^2$ which is always negative (for non zero y) therefore a sattlepoint:

Sattle-points :  (-4,2,-64) and (4,-2,64)

\section{3. Grocery Store}

A grocery store sells two brands of a product, the "house" brand and a "name" brand.  The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell $81 - 21x + 17y$ units of the "house" brand and $40 + 11x - 23y$ units of the "name" brand.

Step 1

Find the revenue function R(x, y).

$R(x,y)=x(81−21x+17y)+y(40+11x−23y) = 81x+40y−21x^2−23y^2+28xy$


Step 2

What is the revenue if she sells the "house" brand for USD 2.30 and the "name" brand for USD 4.10?

```{r }

x = 2.3

y = 4.1

81 * x + 40*y  - 21* x^2  - 23*y^2 + 28 *y*x 

```

\section{4. Cost minimum}

A company has a plant in Los Angeles and a plant in Denver.  The firm is committed to produce a total of 96 units of a product each week.  The total weekly cost is given by $C(x, y) = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700$,  where $x$ is the number of units produced in Los Angeles and $y$ is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

Since we know $x + y = 96$ we can transfor the formula with 2 variables into a formula with 1 substituting y with 96 - x:

$C(x, y) = \frac16x^2 + \frac16 (96-x)^2 + 7x + 25(96-x) + 700$

$= \frac16x^2 + \frac16 (96-x)^2 + 7x + 25(96-x) + 700$

$C'(x, y) = \frac13 x − \frac13(96−x)−18 = \frac23 x −50$ 

Setting this to zero results in x=75

$C''(x, y) =  \frac23$ therefore a minimum

\section{5. Integration}

Evaluate the double integral on the given region.

$\int\int_Re^{(8x + 3y)}dA$ ; $R: 2 \le x \le 4$ and $2\le y  \le 4$


Write your answer in exact form without decimals.

$\int\int_R e^{(8x + 3y)}dA = \int\int_R e^{8x} * e^{3y}dA$

Since the respective integrals are finite we can take them apart:

$\int_2^4 e^{8x} dx \int_2^4e^{3y}dy$

$\frac18 [e^{8x}]_2^4 * \frac13[e^{3y}]_2^4$

$\frac1{24} * (e^{32}-e^{16}) * (e^{12}-e^{6})$

$\frac1{24} * (e^{44}-e^{38} - e^{28}+e^{22})$

Github (both PDF and RMarkdown):

https://github.com/chilleundso/Data605_CompMath/tree/master/Homework15