---
title: "DATA 605 - Homework 7"
author: "Manolis Manoli"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r loading-libraries, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(psych)
library(dplyr)
library(knitr)
library(tidyr)
```


#### 1)

Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y . 

We can think of this as a combinatorical problem: 

In order for $Y_n = y$ we know that at least one $X_i = y$ and all others are great than or equal to y.

Once we know how many ways we can achieve this we can divide by the total number of ways n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k can be chosen (which is $k^n$).

We can start with y=1: This includes all combinations of $X_i$ except those where not a single one has value 1. We can therefore take all possible combinations (again, this is $k^n$) and subtract the amount of combinations when none are 1 ($(k-1)^n$):

$k^n - (k-1)^n$

We can now itteratively go higher with y:

For y=2 we go similarly to the above we start with all possibilitites and subtract all combinations where all $X_i > 2$ (which has $(k-2)^n$ combinations) and obviously we subtract all combinations where Y=1 (which is a disjointed group and we know from above is $k^n - (k-1)^n$):

$k^n - (k-2)^n - ((k^n - (k-1)^n)$

$=-(k-2)^n + (k-1)^n = (k-1)^n -(k-2)^n$

For Y= 3 (all - all $X_i > 3$ - $Y=2$ - $Y=1$):

$k^n - [(k-3)^n] - [(k-1)^n -(k-2)^n] -[k^n - (k-1)^n]$

We see the $k^n$ and $(k-1)^n$ cancel out leaving us with:

$(k-2)^n - (k-3)^n$

The pattern shows that more generally there are the following amount of ways to achieve Y=y:

$(k-y+1)^n - (k-y)^n$

And therefore the probability distribtion is:

$\frac{ (k-y+1)^n - (k-y)^n }{ k^n }$

\newpage



#### 2)

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).  This machine has a manufacturer’s expected lifetime of 10 years.  This means that we expect one failure every ten years.  (Include the probability statements and R Code for each part.). 

##### a.  

What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a geometric.  (Hint: the probability is equivalent to not failing during the first 8 years.) 

Probability to not fail 8 years in a row:
```{r , eval=TRUE, warning=F}
p=1/10

(1-p)^8 
``` 

Using R: pgeom is cumulative i.e. prob that it fails on or before n+1 trial (we want the complement) :
```{r , eval=TRUE, warning=F}
1-pgeom(8-1,0.1)
``` 

Expected value:
```{r , eval=TRUE, warning=F}
E=1/p
E
``` 

standard variation:
```{r , eval=TRUE, warning=F}
Sd=sqrt((1-p)/p^2)
Sd
``` 


##### b.

What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as an exponential.


Probability to not fail 8 years in a row:
```{r , eval=TRUE, warning=F}
l=1/10

exp(-8*.1)
``` 

Using R: pexp is cumulative i.e. prob that it fails on or before n trial (we want the complement) :
```{r , eval=TRUE, warning=F}
1-pexp(8,l)
``` 

Expected value:
```{r , eval=TRUE, warning=F}
E=1/l
E
``` 

standard variation:
```{r , eval=TRUE, warning=F}
Sd=sqrt(1/l^2)
Sd
``` 

##### c.

What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a binomial.  (Hint:  0 success in 8 years)   

Probability to not fail 8 years in a row:
```{r , eval=TRUE, warning=F}
n = 8 
k = 0 
p = 0.1 
q=1-p

choose(n,k) * p^(k) * q^(n-k)

``` 

Using R: dbinom with 8 years with zero breaks:
```{r , eval=TRUE, warning=F}
dbinom(k, n, p)
``` 

Expected value:
```{r , eval=TRUE, warning=F}
E=n*p
E
``` 

standard variation:
```{r , eval=TRUE, warning=F}
Sd=sqrt(n*p*q)
Sd
``` 



##### d.

What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a Poisson
Probability to not fail 8 years in a row:

```{r , eval=TRUE, warning=F}
l <- 8/10
k <- 0

exp(-l)*l^0/factorial(0)

``` 

Using R: ppois:
```{r , eval=TRUE, warning=F}
ppois(k,l)
``` 

Expected value:
```{r , eval=TRUE, warning=F}
E=l
E
``` 

standard variation:
```{r , eval=TRUE, warning=F}
Sd=sqrt(l)
Sd
``` 
 
   
Github (both PDF and RMarkdown):

https://github.com/chilleundso/Data605_CompMath/tree/master/Homework7