---
title: "DATA 605 - Homework 2"
author: "Manolis Manoli"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r loading-libraries, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(psych)
library(dplyr)
library(knitr)
library(tidyr)
```

## Problem Set 1

#### 1)

given a 2 dim A = 

a[1,1]a[1,2]

a[2,1]a[2,2]

the transpose is AT:

a[1,1]a[2,1]

a[1,2]a[2,2]

the first cell of ATA[1,1]

[a1,1] * a[1,1] + a[2,1] * a[2,1]

while first cell of AAT[1,1] is

[a1,1] * a[1,1] + a[1,2] * a[1,2]

clearly if a[1,2] and a[2,1] are different values ATA and AAT will not be the same

for example A:

1 2

3 4

A*AT = 

5 11

11 25


AT * A = 

10 14

14 20

#### 2)

Of course if A is symetrical i.e AT = A then ATA = AAT = AA. However, there are also non-symmetrical matrizes that fulfill ATA = AAT. The full class of these matrices is called **normal**. 

## Problem Set 2

```{r LU-function, eval=TRUE}

LU <- function(A) {
  
  if (dim(A)[1] != dim(A)[2]) {
    notsquare = "not a square matrix"
    return (notsquare)}
  else {
    d = dim(A)
    L =  matrix(0, d, d)
    for (i in 1:d){
      L[i,i]=1
    }
    for (r in 2:d){
      c_end = r-1
      for (c in 1:c_end){
        f = - A[r,c] / A[c,c]
        A[r,]= f * A[c,] + A[r,]
        L[r,c] = -f
        }
      }
    
    }
    return (list(U=A,L=L))
  
}

```

### Test Matrix 1

```{r LU-example1, eval=TRUE, warning=F}

A = matrix(c(1,4,-3,-2,8,5,3,4,7),nrow=3,ncol=3,byrow = T)
LU_A = LU(A)
LU_A$L
LU_A$U

```

### Test Matrix 2

```{r LU-example2, eval=TRUE,  warning=F}

A = matrix(c(2,4,-4,1,-4,3,-6,-9,5),nrow=3,ncol=3,byrow = T)
LU_A = LU(A)
LU_A$L
LU_A$U

```

   
Github (both PDF and RMarkdown):

https://github.com/chilleundso/Data605_CompMath/tree/master/Homework2
    

