---
title: "DATA 605 - Homework 11"
author: "Manolis Manoli"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---
```{r loading-libraries, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(psych)
library(dplyr)
library(knitr)
library(tidyr)
library(GGally)
```
\section{World Health Organisation}

The attached who.csv dataset contains real-world data from 2008.  The variables included follow.

Country:  name of the country

LifeExp:  average life expectancy for the country in years

InfantSurvival:  proportion of those surviving to one year or more

Under5Survival:  proportion of those surviving to five years or more 

TBFree:  proportion of the population without TB.

PropMD:  proportion of the population who are MDs

PropRN:  proportion of the population who are RNs

PersExp:  mean personal expenditures on healthcare in US dollars at average exchange rate 

GovtExp:  mean government expenditures per capita on healthcare, US dollars at average exchange rate

TotExp:  sum of personal and government expenditures.

1.Provide a scatterplot of LifeExp~TotExp, and run simple linear regression.  Do not transform the variables.  Provide and interpret the F statistics, R^2, standard error,and p-values only.  Discuss whether the assumptions of simple linear regression met.

2.Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6).  Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6  as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables.  Provide and interpret the F statistics, R^2, standard error, and p-values.   Which model is "better?"

3.Using the results from 3, forecast life expectancy when TotExp^.06 =1.5.  Then forecast life expectancy when TotExp^.06=2.5.  

4.Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values.  How good is the model?

LifeExp = b0+b1 x PropMd + b2 x TotExp  +b3 x PropMD x TotExp

5.  Forecast LifeExp when PropMD=.03 and TotExp = 14.  Does this forecast seem realistic?  Why or why not?



\section{Solution}



```{r,eval=T}
url = "https://raw.githubusercontent.com/chilleundso/Data605_CompMath/master/Homework12/who.csv"
df = read.csv(url)
```

\subsection{Question 1}

1. Provide a scatterplot of LifeExp~TotExp, and run simple linear regression.  Do not transform the variables.  Provide and interpret the F statistics, R^2, standard error,and p-values only.  Discuss whether the assumptions of simple linear regression met.

```{r,eval=T}
expendVSexpect.lm <- lm(df$LifeExp ~ df$TotExp)
```


```{r,eval=T}
plot(df$LifeExp ~ df$TotExp,main='Personal and government expenditures vs life expectancy', xlab = 'Personal and Government Expenditures', ylab = 'Life Expectancy')
abline(expendVSexpect.lm)

```


```{r,eval=T}
summary(expendVSexpect.lm)
```

Since the p-Value is quite small this model significantly better describes the data than the null hypothesis (average of dependant variable). However looking at the R-squared of 25% we can see that only a quarter of the datas variance is described by the model which is not very satisfying.


\subsection{Question 2}

2. Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6).  Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6  as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables.  Provide and interpret the F statistics, R^2, standard error, and p-values.   Which model is "better?"

```{r,eval=T}
expendVSexpectRAISED.lm <- lm(df$LifeExp^4.6 ~ I(df$TotExp^0.06))
summary(expendVSexpectRAISED.lm)
```

We can see tt the R-Squared has gone up to 73%, now explaining a far larger amount of the variability and the F-statistic has gone from  65 to 508, showing that the model has increased in significance versus using the average observations.


\subsection{Question 3}


3. Using the results from 3, forecast life expectancy when TotExp^.06 =1.5.  Then forecast life expectancy when TotExp^.06=2.5.  

```{r,eval=T}
coef(expendVSexpectRAISED.lm)
intercept=coef(expendVSexpectRAISED.lm)[1]
slope=coef(expendVSexpectRAISED.lm)[2]
```

life expectancy when TotExp^.06 =1.5

```{r,eval=T}
(intercept + slope * 1.5)^(1/4.6)
```

life expectancy when TotExp^.06 =2.5

```{r,eval=T}
(intercept + slope * 2.5)^(1/4.6)
```

\subsection{Question 4}

4. Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values.  How good is the model?

LifeExp = b0+b1 x PropMd + b2 x TotExp  +b3 x PropMD x TotExp


```{r,eval=T}
multi.lm <- lm(df$LifeExp ~ df$PropMD + df$TotExp + df$PropMD * df$TotExp)
summary(multi.lm)
```

R-squared is less the previous model and with ~35% not very satisfying. The p-value for each variable is below the significance threashold and the model as a whol has a relatively high significance with a p-value: < 2.2e-16.

\subsection{Question 5}

5.  Forecast LifeExp when PropMD=.03 and TotExp = 14.  Does this forecast seem realistic?  Why or why not?

```{r,eval=T}
coef(multi.lm)
a=coef(multi.lm)[1]
b1=coef(multi.lm)[2]
b2=coef(multi.lm)[3]
b3=coef(multi.lm)[4]
PropMD=.03 
TotExp = 14
```

```{r,eval=T}
a + (b1 * PropMD) + (b2 * TotExp) + (b3 * PropMD * TotExp)
```

Clearly the model is not a great predicter given that an average life expecancy of 108 years is very unrealistic (currently).

https://github.com/chilleundso/Data605_CompMath/tree/master/Homework12

