---
title: "DATA 605 - Homework 4"
author: "Manolis Manoli"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r loading-libraries, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(psych)
library(dplyr)
library(knitr)
library(tidyr)
```

## Problem Set 1

#### 1)

We start with matrix A

```{r , eval=TRUE, warning=F}
A = matrix(c(1,2,3,-1,0,4),nrow=2,ncol=3,byrow = T)

A
```

We create X = A(AT) and Y = (AT)A:
```{r , eval=TRUE, warning=F}
X = A %*% t(A)
Y = t(A) %*% A

X
Y
```

EigenValue and EigenVector of X and Y:

```{r , eval=TRUE, warning=F}
eigen(X)

eigen(Y)
```

left-singular, singular values, and right-singular vectors of A:

```{r , eval=TRUE, warning=F}
svd(A)
```

The third eigenvalue of Y is actually zero (shows as e-16).

We can see that our manual calculation and the built in svd has the same results.

```{r , eval=TRUE, warning=F}
svd(A)$d**2
```

we can see that the square is equal to the eigenvalue



## Problem Set 2

#### 1)




```{r , eval=TRUE, warning=F}
myinverse <- function(A) {
  

cofactor <- matrix(0,nrow(A),ncol(A))
  for(i in 1:nrow(A)) {
    for(j in 1:ncol(A)) {
      
      temp = A
      temp = temp[-i,] #remove row i
      temp = temp[,-j] #remove col j
      cofactor[i,j] <- (-1)^(i+j) * det(temp)
    }
  }
  
#from the weekly material we know that A-1 = cT / det(A)
  
return(t(cofactor)/ det(A))
} 
```

We will use the below matrix to test the function:

```{r , eval=TRUE, warning=F}
A = matrix(c(1,2,3,4,5,6,7,8,10),nrow=3,ncol=3,byrow = T)

A
```

```{r , eval=TRUE, warning=F}
my_inv_A = myinverse(A)

my_inv_A 
```

```{r , eval=TRUE, warning=F}
inv_A = solve(A)

inv_A 
```

We can see that both inverses are the same

   
Github (both PDF and RMarkdown):

https://github.com/chilleundso/Data605_CompMath/tree/master/Homework4
    

