---
title: "DATA 605 - Homework 11"
author: "Manolis Manoli"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---
```{r loading-libraries, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(psych)
library(dplyr)
library(knitr)
library(tidyr)
library(GGally)
```
\section{Cars - stopping distance vs speed}

Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

\section{Solution}

```{r,eval=T}
data("cars")
```

\subsection{Visualize the Data}

```{r,eval=T}
plot(cars$speed,cars$dist, main="Breaking Distance vs Speed",
xlab="Speed", ylab="Breaking Distance")
```

As expected, there is a general tendency of increasing distance with increasing speed.

\subsection{The Linear Model Function}

```{r,eval=T}
attach(cars)
cars.lm <- lm(dist ~ speed)
cars.lm
```

```{r,eval=T}
plot(speed,dist)
abline(cars.lm)
```


\subsection{Evaluating the Quality of the Model}

```{r,eval=T}
summary(cars.lm)
```

The P-value is very small and R-squared of 65% which in comination would indicate that it's a good fit

\subsection{Residual Analysis}

```{r,eval=T}

 plot(fitted(cars.lm),resid(cars.lm))

```

It seems like there is a little bit of a U shape in the residual. The errors in the middle of the plot seem to be mostly below zero.


```{r,eval=T}
qqnorm(resid(cars.lm))
qqline(resid(cars.lm))
```

We can see that the tails are diverging from the line slightly which might indicate slightly fatter tails than that of a normal distribution.

Overall, I would have assumed a quadratic dependency between speed and distance - however I actualyl do not see this very clearly in the plot we have used so far.



https://github.com/chilleundso/Data605_CompMath/tree/master/Homework11

